\singlespacing \vspace{-1.5cm} \LARGE Trade-offs among plant reproductive traits determine interactions with floral visitors

### 2021-10-19 Manuscript/Manuscript.tex
edits

\textbf{Plant life strategies are constrained by cost-benefit trade-offs
that determine plant form and function. Despite the recent advances in
the understanding of plant trade-offs with vegetative and physiological
traits, little is known about the plant reproductive economics and how
they constrain plant life strategies and shape interactions with floral
visitors. Here, we investigated plant reproductive trade-offs and their
association with floral visitors with a dataset of 1506 plant species
with 17 reproductive and 3 vegetative traits from 28 plant-pollinator
network studies across 18 countries. Then, we tested whether a species
reproductive strategy predicted plant-pollinator interactions
(presence-absence and visitation rate) and if the interaction strategy
(its role in a pollination network) was predicted by individual traits.
We found that 51.8\% of trait variation was explained by two independent
axes that encompassed plant form and function. Specifically, the first
axis indicated the presence of a trade-off between flower number and
flower size (PC1, 26.72\%) and the second axis indicated a trade-off for
the level of pollinator dependency (PC2, 25.08\%). The reproductive
trade-offs explained some of the qualitative interaction with
pollinators (presence-absence) but none of the variability of visitation
rates. However, noteworthy differences across pollinator guilds were
found in both cases. Finally, we found that specific traits with special
relevance of floral rewards were key in the understanding of the species
network role. Our results highlight the need to consider plant
reproductive trade-offs to improve our understanding of plant life
strategies and plant-pollinator interactions at broader spatial scales.}

@@

Despite the astonishing diversity of floral structures among flowering
plants\textsuperscript{\protect\hyperlink{ref-barrett2002}{1},\protect\hyperlink{ref-schiestl2013}{2}}
and their relevance to understand plant-pollinator
interactions\textsuperscript{\protect\hyperlink{ref-fenster2004}{3},\protect\hyperlink{ref-dellinger2020}{4}},
a unified framework that explores plant reproductive compromises is
currently lacking\textsuperscript{\protect\hyperlink{ref-roddy2021}{5}}.
In addition, macroecological studies regarding specific reproductive
traits are
scarce\textsuperscript{\protect\hyperlink{ref-baude2016}{6}--\protect\hyperlink{ref-moeller2017}{9}}
and as a consequence studies that link reproductive traits to
plant-pollinator interactions at large scales are poorly
explored\textsuperscript{\protect\hyperlink{ref-sargent2008}{10}--\protect\hyperlink{ref-ruger2018}{13}}.
Linking the position in the trait-space with the different pollinator
groups could help to improve our understanding of plant-pollinator
associations\textsuperscript{\protect\hyperlink{ref-dehling2016}{14}}.
At the same time, there is increasing interest in the understanding of
the determinants of plant-pollinator interactions via trait-based
approaches\textsuperscript{\protect\hyperlink{ref-fenster2004}{3},\protect\hyperlink{ref-rosas2014}{15}}
and trait-matching
analyses\textsuperscript{\protect\hyperlink{ref-stang2009}{16},\protect\hyperlink{ref-bartomeus2016}{17}}.
However, reproductive traits have been overlooked beyond highly
specialised
systems\textsuperscript{\protect\hyperlink{ref-dellinger2020}{4}}
despite the apparent generalist nature of plant-pollinator
interactions\textsuperscript{\protect\hyperlink{ref-waser1996}{18},\protect\hyperlink{ref-olesen2002}{19}}.
In addition, it is yet unclear how the different reproductive biology
traits influence plant-pollinator
associations\textsuperscript{\protect\hyperlink{ref-tur2013}{20},\protect\hyperlink{ref-devaux2014}{21}}.

Species can optimize their fitness through various life-history traits,
yet trade-offs among those traits constrain the range of potential
strategies that a species can employ. With the recent availability of
large trait databases (e.g.,
TRY\textsuperscript{\protect\hyperlink{ref-kattge2011}{22}} and
COMPADRE\textsuperscript{\protect\hyperlink{ref-salguero2015}{23}}),
plant ecological strategies are increasingly being examined, and are
facilitating the identification of global patterns and constraints of
plant form and
function\textsuperscript{\protect\hyperlink{ref-salguero2016}{12},\protect\hyperlink{ref-diaz2016}{24},\protect\hyperlink{ref-carmona2021}{25}}.
However, the main focus has been on vegetative traits such as
leaf\textsuperscript{\protect\hyperlink{ref-wright2004}{26}},
wood\textsuperscript{\protect\hyperlink{ref-chave2009}{27}}, or
root\textsuperscript{\protect\hyperlink{ref-laughlin2021}{28}}
trade-offs with little or no attention given to reproductive
traits\textsuperscript{\protect\hyperlink{ref-roddy2021}{5},\protect\hyperlink{ref-evojtko2020}{29}},
also critical to plant life strategies that shape plant interactions
with pollinators and ultimately impact plant reproductive success. For
instance, short lived versus perennial species tend to have low versus
high levels of outcrossing,
respectively\textsuperscript{\protect\hyperlink{ref-moeller2017}{9},\protect\hyperlink{ref-barrett2003}{30}}.
Further, outcrossing levels are positively correlated with flower
size\textsuperscript{\protect\hyperlink{ref-goodwillie2010}{31}}. In
addition, the presence of costly rewards (e.g., pollen or nectar) and
showy flowers or floral displays can only be understood through
consideration of plant species' reliance upon animal pollination
(pollinator dependence) and their role in attracting
pollinators\textsuperscript{\protect\hyperlink{ref-ollerton2011}{32}}.
However, it is still unknown to what extent these different reproductive
compromises determine plant-pollinator associations.

Several studies have identified links between plant traits and
plant-pollinator network
properties\textsuperscript{\protect\hyperlink{ref-bartomeus2013}{33}--\protect\hyperlink{ref-rowe2020}{35}}.
Moreover, plant traits can also define species' network
roles\textsuperscript{\protect\hyperlink{ref-tur2013}{20},\protect\hyperlink{ref-lazaro2013}{36}}
(e.g., specialists vs generalists). For example, species that occupy
trait space extremes are more likely to exhibit higher specialization
and be more reliant on
trait-matching\textsuperscript{\protect\hyperlink{ref-junker2013}{37},\protect\hyperlink{ref-coux2016}{38}}.
This morphological matching between plant and floral visitors can
determine plant-pollinator interactions, and thus shape their
interaction network
structure\textsuperscript{\protect\hyperlink{ref-stang2009}{16},\protect\hyperlink{ref-ibanez2012}{39}}.
Remarkably, the combination of traits have shown to increase the
predictive power of the network
interactions\textsuperscript{\protect\hyperlink{ref-eklof2013}{40}}.
Therefore, considering the different plant reproductive trade-offs which
represent the species strategy within the
network\textsuperscript{\protect\hyperlink{ref-dehling2016}{14}} could
progress our understanding of plant-pollinator interactions. Further, we
know little if the different network patterns generally studied at the
community level are maintained when we scale up at macroecological
scale.

Here, we explore the potential trade-offs among plant floral and
reproductive biology traits and how these influence plant-pollinator
interactions. First, we identify the major axes of reproductive trait
variation and trade-offs that determine plant form and function. Second,
we investigate how plant species' position in trait-space influences
plant-pollinator interactions (presence-absence and visitation rate).
Finally, we investigate how the main axes of trait variation and
individual traits influence plant species roles within networks using
complementary interaction network metrics (i.e., interaction strength,
normalized degree and specialization).

@@

\textbf{Plant strategies.} The phylogenetically informed principal
component analysis (pPCA) captured by the first two and three axes
51.8\% and 70.97\% of trait variation, respectively (Fig. 1 and
Supplementary Fig. S5) and had a phylogenetic correlation (\(\lambda\))
of 0.76. The first principal component (PC1) represented 26.72\% of the
trait variation and indicated a trade-off between flower number and
flower size. We refer to this axis as the `flower number - flower size
trade-off', as already described in previous
studies\textsuperscript{\protect\hyperlink{ref-sargent2007}{41},\protect\hyperlink{ref-kettle2011}{42}}.
Hence, one end of the spectrum comprised species with high investment in
flower number and plant height but small flower size, short style length
and low ovule number. The other end of this spectrum comprised species
that were short in height and invested in large flowers, long styles,
many ovules, but few flowers. The main contributing traits to PC1 were
plant height, flower number, ovule number and flower size (loadings
\textgreater{} \textbar{}0.5\textbar{}; Supplementary Table S3) but
style length also contributed moderately on PC1 (loading = -0.33). The
second principal component (PC2) represented 25.05\% of the trait
variation and indicated a trade-off between low and high pollinator
dependence. We refer to this axis as the `pollinator dependence
trade-off'. The main driver of trait variation on PC2 was autonomous
selfing (loading = 0.85) but the other traits (except ovule number) also
made moderate contributions (loadings from 0.27 to 0.4; Supplementary
Table S3). We found that high pollinator dependence was associated with
larger and a higher number of flowers, greater plant height and longer
styles. In contrast, species with high levels of autonomous selfing
tended to have fewer and smaller flowers, had shorter styles and were
shorter in height. Further, PC3 explained a considerable amount of trait
variability (19.17\%) and the main contributors to this axis were style
length (loading = -0.66) and the degree of autonomous selfing (loading =
-0.51). The remaining traits, apart from ovule number, were moderately
correlated to changes on PC3 (loadings from -0.23 to -0.46;
Supplementary Table S3). Thus, because style length was correlated with
all traits on PC3 and was the main driver of trait variation, we refer
to this axis as the `style length trade-off'. Further, the pPCA with the
subset of species that had nectar and pollen quantity data showed that
nectar quantity (microlitres of nectar per flower) was positively
associated with flower size, style length and ovule number (PC1,
23.40\%); and pollen quantity (pollen grains per flower) was positively
correlated with flower number and plant height and negatively associated
with autonomous selfing (PC2, 21.67\%; Supplementary Fig. S6). This pPCA
explained similar variance with the first two principal components
(45.07\%) and similar associations of traits despite some variability in
the loadings (Supplementary Table S4).

@@

Here, we show that plant species exhibit clear trade-offs in their
floral, reproductive biology and vegetative traits. These trade-offs are
differentiated on three main axes of trait variation: (i) flower number
- flower size, (ii) pollinator dependence and (iii) style length. These
axes of trait variation explained partly the presence-absence
interactions between plants and floral visitors but did not explain the
overall visitation rates. However, on both qualitative and quantitative
interactions floral visitor guilds showed associations with the
different reproductive trade-offs. Interestingly, pollen and nectar
related traits were better than all other traits for characterizing
plant species functional roles within their interaction networks.

Over 50\% of plant trait variation was captured by the flower number -
flower size and pollinator dependence trade-offs. Variation in these two
axes of trait variation appears to be associated with the commonly
described `fast-slow continuum' in
plant\textsuperscript{\protect\hyperlink{ref-salguero2016}{12}} and
animal\textsuperscript{\protect\hyperlink{ref-healy2019}{43}}
life-history strategies as indicated by the different floral and
reproductive biology traits associated with plant height, life form and
lifespan. The `slow' part of the continuum (i.e., tall trees and shrubs)
had plant species with many flowers, few ovules, higher pollinator
dependence, greater occurrence of self-incompatibility and more complex
breeding systems (e.g., monoecious and dioecious species). In contrast,
species that employed the `fast' strategy (i.e., short herbs), had fewer
flowers, more ovules, greater occurrence of self compatibility and lower
pollinator dependence. Further, on the first two axes of trait
variation, we found additional support for the positive association
already described in a previous study between outcrossing rate and
floral
display\textsuperscript{\protect\hyperlink{ref-goodwillie2010}{31}}.
This trend was further confirmed when looking at the specific
correlation between floral display and autonomous selfing for our
dataset (see Supplementary Fig. S10).

Despite the low predictive power of the main axes of trait variation on
overall interaction patterns (presence-absence and visitation rate), we
found commonalities among and within floral visitors guilds across these
axes that indicate the influence of life-history strategies on
plant-pollinator interactions. For example, all guilds showed an
increasing trend of visitation on plant species with higher pollinator
dependence, which were also associated with larger floral displays and
greater pollen quantities (Fig. 1 and Supplementary Fig. S6). This trend
is consistent with previous plant-pollinator studies that show higher
visitation rates on species that make greater reproductive
investment\textsuperscript{\protect\hyperlink{ref-lazaro2013}{36},\protect\hyperlink{ref-hegland2005}{44},\protect\hyperlink{ref-kaiser2014}{45}}.
In regard to the flower number - flower size and style length
trade-offs, different guilds showed contrasting visitation rates, which
could be associated with different pollination syndromes at a
macroecological scale. For instance, the guilds of bees and syrphids
were clearly associated with opposing life-strategies on PC1 and PC3
(Fig. 3) indicating a possible niche partitioning among these two
guilds\textsuperscript{\protect\hyperlink{ref-palmer2003}{46},\protect\hyperlink{ref-phillips2020}{47}}.
However, these plant-pollinator associations do not account for some of
the traits (i.e., pollen and nectar quantity) that had greater relevance
on the species network roles (Fig. 4) because of insufficient data
availability and lack of pollinator efficiency measurements for more
complete descriptions of the pollination
syndromes\textsuperscript{\protect\hyperlink{ref-dellinger2020}{4}}. In
any case, it is worth noting that this general pattern emerges in
field-measured plant-pollinator networks, where other local factors such
as species relative abundances, will surely explain part of the observed
variability\textsuperscript{\protect\hyperlink{ref-bartomeus2016}{17},\protect\hyperlink{ref-vazquez2007}{48},\protect\hyperlink{ref-encinas2012}{49}}.

To conclude, we provide the first description of plant floral and
reproductive biology trade-offs using a large global dataset of plant
traits and interactions with floral visitors. This allowed us to
identify the major reproductive strategies of flowering plants and how
plant life strategies determine their interaction with floral visitor
guilds. Although the percentage of explained variation found with the
first two axes (\textasciitilde{}50\%) is lower than other studies that
have evaluated life histories with vegetative
traits\textsuperscript{\protect\hyperlink{ref-diaz2016}{24},\protect\hyperlink{ref-carmona2021}{25}},
this percentage is consistent with the largest and most recent study
that has characterized life strategies with both vegetative and
reproductive (i.e., semelparity and iteroparity)
traits\textsuperscript{\protect\hyperlink{ref-salguero2016}{12}}. Future
work needs to integrate these floral and reproductive compromises with
the already described vegetative and physiological
trade-offs\textsuperscript{\protect\hyperlink{ref-diaz2016}{24}} to
create a more comprehensive spectrum of trait
variation\textsuperscript{\protect\hyperlink{ref-roddy2021}{5}}.
Further, the varying level of phylogenetic signal among traits deserves
further attention to understand evolutionary changes on mating and
flower morphology in response to
pollinators\textsuperscript{\protect\hyperlink{ref-gervasi2017}{50},\protect\hyperlink{ref-mackin2021}{51}}.
Finally, a more complete description of the macroecological patterns of
plant-pollinator interactions including unrepresented areas of the
world\textsuperscript{\protect\hyperlink{ref-poisot2021}{52}} and traits
that have been largely ignored (e.g., floral rewards and plant mating
system) are needed to progress in our understanding of the global
patterns in plant-pollinator interactions.

@@

\hypertarget{ref-bartomeus2013}{}
33. Bartomeus, I. Understanding Linkage Rules in Plant-Pollinator
Networks by Using Hierarchical Models That Incorporate Pollinator
Detectability and Plant Traits. \emph{PLoS One} \textbf{8}, e69200
(2013).

\hypertarget{ref-olito2015}{}
34. Olito, C. \& Fox, J. W. Species traits and abundances predict
metrics of plantPollinator network structure, but not pairwise
interactions. \emph{Oikos} \textbf{124}, 428--436 (2015).

\hypertarget{ref-rowe2020}{}
35. Rowe, L. \emph{et al.} Flower traits associated with the visitation
patterns of bees. \emph{Oecologia} \textbf{193}, 511--522 (2020).

\hypertarget{ref-lazaro2013}{}
36. Lázaro, A., Jakobsson, A. \& Totland, Ø. How do pollinator
visitation rate and seed set relate to species' floral traits and
community context? \emph{Oecologia} \textbf{173}, 881--893 (2013).

\hypertarget{ref-junker2013}{}
37. Junker, R. R. \emph{et al.} Specialization on traits as basis for
the niche-breadth of flower visitors and as structuring mechanism of
ecological networks. \emph{Funct. Ecol.} \textbf{27}, 329--341 (2013).

\hypertarget{ref-coux2016}{}
38. Coux, C., Rader, R., Bartomeus, I. \& Tylianakis, J. M. Linking
species functional roles to their network roles. \emph{Ecol. Lett.}
\textbf{19}, 762--770 (2016).

\hypertarget{ref-ibanez2012}{}
39. Ibanez, S. Optimizing size thresholds in a plant-pollinator
interaction web: Towards a mechanistic understanding of ecological
networks. \emph{Oecologia} \textbf{170}, 233--242 (2012).

\hypertarget{ref-eklof2013}{}
40. Eklöf, A. \emph{et al.} The dimensionality of ecological networks.
\emph{Ecol. Lett.} \textbf{16}, 577--583 (2013).

\hypertarget{ref-sargent2007}{}
41. Sargent, R. D., Goodwillie, C., Kalisz, S. \& Ree, R. H.
Phylogenetic evidence for a flower size and number trade-off. \emph{Am.
J. Bot.} \textbf{94}, 2059--2062 (2007).

\hypertarget{ref-kettle2011}{}
42. Kettle, C. J. \emph{et al.} Ecological Implications of a Flower
Size/Number Trade-Off in Tropical Forest Trees. \emph{PLoS One}
\textbf{6}, e16111 (2011).

\hypertarget{ref-healy2019}{}
43. Healy, K., Ezard, T. H. G., Jones, O. R., Salguero-Gómez, R. \&
Buckley, Y. M. Animal life history is shaped by the pace of life and the
distribution of age-specific mortality and reproduction. \emph{Nat.
Ecol. Evol.} \textbf{3}, 1217--1224 (2019).

\hypertarget{ref-hegland2005}{}
44. Hegland, S. J. \& Totland, Ø. Relationships between species' floral
traits and pollinator visitation in a temperate grassland.
\emph{Oecologia} \textbf{145}, 586--594 (2005).

\hypertarget{ref-kaiser2014}{}
45. Kaiser-Bunbury, C. N., Vázquez, D. P., Stang, M. \& Ghazoul, J.
Determinants of the microstructure of plantPollinator networks.
\emph{Ecology} \textbf{95}, 3314--3324 (2014).

>>>>

\textbf{Plant life strategies are constrained by cost-benefit trade-offs
that determine plant form and function. However, despite recent advances
in the understanding of trade-offs for vegetative and physiological
traits, little is known about plant reproductive economics and how they
constrain plant life strategies and shape interactions with floral
visitors. Here, we investigate plant reproductive trade-offs and how
these reproductive compromises drive interactions with floral visitors
using a dataset of 17 reproductive traits for 1,506 plant species from
28 plant-pollinator studies across 18 countries. We tested whether a
plant's reproductive strategy predicts its interactions with floral
visitors and if the different reproductive traits predict the plant's
role within the pollination network. We found that over half of all
plant reproductive trait variation was explained by two independent axes
that encompassed plant form and function. Specifically, the first axis
indicated the presence of a trade-off between flower number and flower
size, while the second axis indicated a pollinator dependency trade-off.
Plant reproductive trade-offs somewhat explained the presence or absence
of interactions with floral visitors, but not differences in visitation
rate. However, we did find important differences in the interaction
level among floral visitor guilds on the different axes of trait
variation. Finally, we found that plant size and floral rewards were the
most important traits in the understanding of the species network role.
Our results highlight the importance of plant reproductive trade-offs in
determining plant life strategies and plant-pollinator interactions in a
global context.}

@@

Despite the astonishing diversity of floral structures among flowering
plants\textsuperscript{\protect\hyperlink{ref-barrett2002}{1},\protect\hyperlink{ref-schiestl2013}{2}}
and their importance in shaping plant-pollinator
interactions\textsuperscript{\protect\hyperlink{ref-fenster2004}{3},\protect\hyperlink{ref-dellinger2020}{4}},
a unified framework that explores plant reproductive compromises is
currently lacking\textsuperscript{\protect\hyperlink{ref-roddy2021}{5}}.
In addition, macroecological studies that investigate plant reproductive
traits are
scarce\textsuperscript{\protect\hyperlink{ref-baude2016}{6}--\protect\hyperlink{ref-moeller2017}{9}}
and consequently, there is poor understanding of how reproductive traits
drive interactions with floral visitors at large
scales\textsuperscript{\protect\hyperlink{ref-sargent2008}{10}--\protect\hyperlink{ref-ruger2018}{13}}.
Linking the position in the trait-space with the different pollinator
groups could help to improve our understanding of plant-pollinator
associations\textsuperscript{\protect\hyperlink{ref-dehling2016}{14}}.
Further, there is increasing interest in understanding drivers of
plant-pollinator interactions using trait-based
approaches\textsuperscript{\protect\hyperlink{ref-fenster2004}{3},\protect\hyperlink{ref-rosas2014}{15}}
and trait-matching
analyses\textsuperscript{\protect\hyperlink{ref-stang2009}{16},\protect\hyperlink{ref-bartomeus2016}{17}}.
However, reproductive traits have been overlooked beyond highly
specialised pollination
systems\textsuperscript{\protect\hyperlink{ref-dellinger2020}{4}}
despite the apparent generalist nature of plant-pollinator
interactions\textsuperscript{\protect\hyperlink{ref-waser1996}{18},\protect\hyperlink{ref-olesen2002}{19}}.
In addition, it is unclear how specific plant reproductive biology
traits influence plant-pollinator
interactions\textsuperscript{\protect\hyperlink{ref-tur2013}{20},\protect\hyperlink{ref-devaux2014}{21}}.

Species can optimise their fitness through various life-history traits,
yet trade-offs among those traits constrain the range of potential
strategies that a species can use. With the recent availability of large
trait databases (e.g.,
TRY\textsuperscript{\protect\hyperlink{ref-kattge2011}{22}} and
COMPADRE\textsuperscript{\protect\hyperlink{ref-salguero2015}{23}}),
plant ecological strategies are being increasingly examined, and are
facilitating the identification of global patterns and constraints in
plant form and
function\textsuperscript{\protect\hyperlink{ref-salguero2016}{12},\protect\hyperlink{ref-diaz2016}{24},\protect\hyperlink{ref-carmona2021}{25}}.
However, most studies have focused on vegetative traits such as
leaf\textsuperscript{\protect\hyperlink{ref-wright2004}{26}},
wood\textsuperscript{\protect\hyperlink{ref-chave2009}{27}}, or
root\textsuperscript{\protect\hyperlink{ref-laughlin2021}{28}}
trade-offs with little or no attention given to reproductive
traits\textsuperscript{\protect\hyperlink{ref-roddy2021}{5},\protect\hyperlink{ref-evojtko2020}{29}}
which are critical to plant life strategies that shape interactions with
pollinators and ultimately determine plant reproductive success. For
instance, short lived versus perennial species tend to have low versus
high levels of outcrossing,
respectively\textsuperscript{\protect\hyperlink{ref-moeller2017}{9},\protect\hyperlink{ref-barrett2003}{30}}
and outcrossing levels are positively correlated with flower
size\textsuperscript{\protect\hyperlink{ref-goodwillie2010}{31}}. In
addition, the presence of costly rewards (e.g., pollen or nectar) and
showy flowers or floral displays can only be understood through
consideration of plant species' reliance upon animal pollination
(pollinator dependence) and its role in attracting
pollinators\textsuperscript{\protect\hyperlink{ref-ollerton2011}{32},\protect\hyperlink{ref-rodger2021}{33}}.
However, it is still unknown to what extent these different reproductive
compromises determine plant-pollinator interactions.

Several studies have identified links between plant traits and
plant-pollinator network
properties\textsuperscript{\protect\hyperlink{ref-bartomeus2013}{34}--\protect\hyperlink{ref-rowe2020}{36}}.
Moreover, plant traits can define species' network roles (e.g.,
specialists vs
generalists)\textsuperscript{\protect\hyperlink{ref-tur2013}{20},\protect\hyperlink{ref-lazaro2013}{37}}.
For example, plant species that occupy reproductive trait space extremes
are more likely to exhibit higher levels of specialisation and be more
reliant on the trait-matching with
pollinators\textsuperscript{\protect\hyperlink{ref-junker2013}{38},\protect\hyperlink{ref-coux2016}{39}}.
Morphological matching between plant and floral visitors often
determines plant-pollinator interactions, and can thus strongly
influence interaction network
structure\textsuperscript{\protect\hyperlink{ref-stang2009}{16},\protect\hyperlink{ref-ibanez2012}{40}}.
Remarkably, the combination of traits have shown to increase the
predictive power of the network
interactions\textsuperscript{\protect\hyperlink{ref-eklof2013}{41}}.
Therefore, considering the different plant reproductive trade-offs which
represent the species strategy within the
network\textsuperscript{\protect\hyperlink{ref-dehling2016}{14}} could
progress our understanding of plant-pollinator interactions. Further, we
know little if the different network patterns generally studied at the
community level are maintained when we scale up at macroecological
scale.

Here, we aim to explore the potential trade-offs among reproductive
traits and how these influence plant-pollinator interactions. First, we
identify the major axes of reproductive trait variation and trade-offs
that determine plant form and function. Second, we investigate how plant
species' position in trait-space interactions with floral visitors.
Finally, we investigate how both the main axes of trait variation, and
individual traits, influence plant species' roles within networks using
a set of complementary interaction network metrics (i.e., interaction
strength, normalized degree and specialization).

@@

\textbf{Plant strategies.} The phylogenetically informed principal
component analysis (pPCA) captured by the first two and three axes
51.8\% and 70.97\% of trait variation, respectively (Fig. 1 and
Supplementary Fig. S5) and had a phylogenetic correlation (\(\lambda\))
of 0.76. The first principal component (PC1) represented 26.72\% of the
trait variation and indicated a trade-off between flower number and
flower size. We refer to this axis as the `flower number - flower size
trade-off', as already described in previous
studies\textsuperscript{\protect\hyperlink{ref-sargent2007}{42},\protect\hyperlink{ref-kettle2011}{43}}.
Hence, one end of the spectrum comprised species with high investment in
flower number and plant height but small flower size, short style length
and low ovule number. The other end of this spectrum comprised species
that were short in height and invested in large flowers, long styles,
many ovules, but few flowers. The main contributing traits to PC1 were
plant height, flower number, ovule number and flower size (loadings
\textgreater{} \textbar{}0.5\textbar{}; Supplementary Table S3) but
style length also contributed moderately on PC1 (loading = -0.33). The
second principal component (PC2) represented 25.05\% of the trait
variation and indicated a trade-off between low and high pollinator
dependence. We refer to this axis as the `pollinator dependence
trade-off'. The main driver of trait variation on PC2 was autonomous
selfing (loading = 0.85) but the other traits (except ovule number) also
made moderate contributions (loadings from 0.27 to 0.4; Supplementary
Table S3). We found that high pollinator dependence was associated with
larger and a higher number of flowers, greater plant height and longer
styles. In contrast, species with high levels of autonomous selfing
tended to have fewer and smaller flowers, had shorter styles and were
shorter in height. Further, PC3 explained a considerable amount of trait
variability (19.17\%) and the main contributors to this axis were style
length (loading = -0.66) and the degree of autonomous selfing (loading =
-0.51). The remaining traits, apart from ovule number, were moderately
correlated to changes on PC3 (loadings from -0.23 to -0.46;
Supplementary Table S3). Thus, because style length was correlated with
all traits on PC3 and was the main driver of trait variation, we refer
to this axis as the `style length trade-off'. Further, the pPCA with the
subset of species that had nectar and pollen quantity data showed that
nectar quantity (microlitres of nectar per flower) was positively
associated with flower size, style length and ovule number (PC1,
23.40\%); and pollen quantity (pollen grains per flower) was positively
correlated with flower number and plant height and negatively associated
with autonomous selfing (PC2, 21.67\%; Supplementary Fig. S6). This pPCA
explained similar variance with the first two principal components
(45.07\%) and similar associations of traits despite some variability in
the loadings (Supplementary Table S4).

@@

Here, we show that plant species exhibit clear trade-offs among their
vegetative and reproductive traits that determined different interaction
levels with floral visitor guilds. These trade-offs are differentiated
on three main axes of trait variation: (i) flower number - flower size,
(ii) pollinator dependence and (iii) style length. These reproductive
trade-offs somewhat explained the presence or absence of interactions
with floral visitors, but not differences in visitation rate. However,
on both qualitative and quantitative interactions we found marked
associations of floral visitor guilds with the main axes of trait
variation. Further, we found that plant height and floral reward related
traits were best at characterising plant species functional roles within
pollination networks.

Over half of all plant trait variation was captured by the flower number
- flower size and pollinator dependence trade-offs. Trait variation on
these two axes was associated with the `fast-slow continuum' in
plant\textsuperscript{\protect\hyperlink{ref-salguero2016}{12}} and
animal\textsuperscript{\protect\hyperlink{ref-healy2019}{44}}
life-history strategies, as indicated by the different floral and
reproductive biology traits associated with plant height, life form and
lifespan. The `slow' part of this continuum (i.e., tall trees and
shrubs) included plant species with many flowers, few ovules, higher
pollinator dependence, frequent occurrence of self-incompatibility and
more complex breeding systems (e.g., monoecious and dioecious species).
In contrast, plant species that employed the `fast' strategy (i.e.,
short herbs), had fewer flowers, more ovules, frequent occurrence of
self-compatibility and lower pollinator dependence. Further, on the
first two axes of trait variation, we found additional support for the
previously described positive association between higher outcrossing
rate and larger floral
display\textsuperscript{\protect\hyperlink{ref-goodwillie2010}{31}}. The
positive correlation between larger floral display and higher pollinator
dependence in our dataset further confirmed this trend (see
Supplementary Fig. S10).

Despite the low predictive power of the main trait variation axes for
broad-level interaction patterns (presence-absence of interactions and
visitation rate), we found changes in the interaction patterns among and
within floral visitor guilds across these axes that suggest plant
life-history strategies strongly influence plant-pollinator
interactions. For example, all floral visitor guilds visited plant
species with higher pollinator dependence more frequently, and high
pollinator dependence was associated with large floral displays and
greater pollen quantities (Fig. 1 and Supplementary Fig. S6). This trend
is consistent with previous studies that show plant species with higher
reproductive investment tend to be visited by pollinators more
frequently\textsuperscript{\protect\hyperlink{ref-lazaro2013}{37},\protect\hyperlink{ref-hegland2005}{45},\protect\hyperlink{ref-kaiser2014}{46}}.
In regard to the flower number - flower size and style length
trade-offs, different pollinator guilds showed contrasting visitation
rates across the continuum of trait variation, which could be associated
with different pollination syndromes at a macroecological scale. For
instance, bees and syrphid flies were clearly associated with opposing
life-strategies on PC1 and PC3 (Fig. 3) suggesting possible niche
partitioning\textsuperscript{\protect\hyperlink{ref-palmer2003}{47},\protect\hyperlink{ref-phillips2020}{48}}
between these two guilds. However, despite floral rewards, related
traits were among the best at characterising species functional roles
(Fig. 4), yet the association between reproductive trade-offs and floral
visitors did not account for floral rewards because there was
insufficient data available. More detailed exploration of reproductive
trade-offs in conjunction with floral rewards is needed to help
elucidate plant-pollinator associations. In any case, it is worth noting
that other local factors such as species relative abundances, surely
explain part of the observed
variability\textsuperscript{\protect\hyperlink{ref-bartomeus2016}{17},\protect\hyperlink{ref-vazquez2007}{49},\protect\hyperlink{ref-encinas2012}{50}}
that reproductive trade-offs do not.

To conclude, we provide the first description of plant reproductive
trade-offs using a large global dataset of plant traits. We identified
the major reproductive strategies of flowering plants and how these
strategies influence interactions with different floral visitor guilds.
Although the explained variation that we found in the first two axes is
lower than previous studies of vegetative
traits\textsuperscript{\protect\hyperlink{ref-diaz2016}{24},\protect\hyperlink{ref-carmona2021}{25}}
it is consistent with the largest and most recent study that has
characterised plant life strategies with vegetative and reproductive
traits\textsuperscript{\protect\hyperlink{ref-salguero2016}{12}}. Future
work needs to integrate the reproductive compromises that we have
identified with vegetative and physiological trade-offs to create a more
comprehensive spectrum of plant trait variation. Further, the varying
level of phylogenetic signal among traits deserves further attention to
understand evolutionary changes on mating and flower morphology in
response to
pollinators\textsuperscript{\protect\hyperlink{ref-gervasi2017}{51},\protect\hyperlink{ref-mackin2021}{52}}.
Finally, including plant-pollinator networks from unrepresented areas of
the world and a more complete description of plant reproductive
trade-offs is essential for a better understanding of the global
patterns in plant-pollinator interactions.

@@

\hypertarget{ref-rodger2021}{}
33. Rodger, J. G. \emph{et al.} Widespread vulnerability of flowering
plant seed production to pollinator declines. \emph{Science Advances}
\textbf{7}, eabd3524 (2021).

\hypertarget{ref-bartomeus2013}{}
34. Bartomeus, I. Understanding Linkage Rules in Plant-Pollinator
Networks by Using Hierarchical Models That Incorporate Pollinator
Detectability and Plant Traits. \emph{PLoS One} \textbf{8}, e69200
(2013).

\hypertarget{ref-olito2015}{}
35. Olito, C. \& Fox, J. W. Species traits and abundances predict
metrics of plantPollinator network structure, but not pairwise
interactions. \emph{Oikos} \textbf{124}, 428--436 (2015).

\hypertarget{ref-rowe2020}{}
36. Rowe, L. \emph{et al.} Flower traits associated with the visitation
patterns of bees. \emph{Oecologia} \textbf{193}, 511--522 (2020).

\hypertarget{ref-lazaro2013}{}
37. Lázaro, A., Jakobsson, A. \& Totland, Ø. How do pollinator
visitation rate and seed set relate to species' floral traits and
community context? \emph{Oecologia} \textbf{173}, 881--893 (2013).

\hypertarget{ref-junker2013}{}
38. Junker, R. R. \emph{et al.} Specialization on traits as basis for
the niche-breadth of flower visitors and as structuring mechanism of
ecological networks. \emph{Funct. Ecol.} \textbf{27}, 329--341 (2013).

\hypertarget{ref-coux2016}{}
39. Coux, C., Rader, R., Bartomeus, I. \& Tylianakis, J. M. Linking
species functional roles to their network roles. \emph{Ecol. Lett.}
\textbf{19}, 762--770 (2016).

\hypertarget{ref-ibanez2012}{}
40. Ibanez, S. Optimizing size thresholds in a plant-pollinator
interaction web: Towards a mechanistic understanding of ecological
networks. \emph{Oecologia} \textbf{170}, 233--242 (2012).

\hypertarget{ref-eklof2013}{}
41. Eklöf, A. \emph{et al.} The dimensionality of ecological networks.
\emph{Ecol. Lett.} \textbf{16}, 577--583 (2013).

\hypertarget{ref-sargent2007}{}
42. Sargent, R. D., Goodwillie, C., Kalisz, S. \& Ree, R. H.
Phylogenetic evidence for a flower size and number trade-off. \emph{Am.
J. Bot.} \textbf{94}, 2059--2062 (2007).

\hypertarget{ref-kettle2011}{}
43. Kettle, C. J. \emph{et al.} Ecological Implications of a Flower
Size/Number Trade-Off in Tropical Forest Trees. \emph{PLoS One}
\textbf{6}, e16111 (2011).

\hypertarget{ref-healy2019}{}
44. Healy, K., Ezard, T. H. G., Jones, O. R., Salguero-Gómez, R. \&
Buckley, Y. M. Animal life history is shaped by the pace of life and the
distribution of age-specific mortality and reproduction. \emph{Nat.
Ecol. Evol.} \textbf{3}, 1217--1224 (2019).

\hypertarget{ref-hegland2005}{}
45. Hegland, S. J. \& Totland, Ø. Relationships between species' floral
traits and pollinator visitation in a temperate grassland.
\emph{Oecologia} \textbf{145}, 586--594 (2005).

\hypertarget{ref-kaiser2014}{}
46. Kaiser-Bunbury, C. N., Vázquez, D. P., Stang, M. \& Ghazoul, J.
Determinants of the microstructure of plantPollinator networks.
\emph{Ecology} \textbf{95}, 3314--3324 (2014).
